### 2018-07-05 paper/2-background.tex
Address suggestions of reviewers
Tweaks to S2

Build systems automate the execution of repeatable tasks for individual users
and large organisations. In this section we explore the design space of build
systems, using four concrete examples: \Make~\cite{feldman1979make},
\Shake~\cite{mitchell2012shake}, \Bazel~\cite{bazel}, and
\Excel~\cite{advanced_excel}\footnote{\Excel appears very different to the
others but, seen through the lens of this paper, it is very close indeed.}.
We have carefully chosen these four to illustrate the various axes on which
build systems differ; we discuss many other notable examples of build systems,
and their relationships, in~\S\ref{sec-related}.

@@

\noindent
There are two input cells \cmd{A1} and \cmd{A2}, and a single task that computes
the sum of their values, writing the result into the cell \cmd{B1}. If either of
the inputs change, \Excel will recompute the result.

@@

\noindent
The formula in \cmd{B1} uses the \cmd{INDIRECT} function, which takes a string
and returns the value of the cell with that name.  The string evaluates to
\cmd{"A1"}, so \cmd{B1} evaluates to \cmd{10}. However, the dependencies of the
formula in \cmd{B1} are determined by the value of \cmd{C1}, so there is no
possibility\footnote{In this particular example one might say that the value of
\cmd{C1} is statically known, but imagine that it is the result of a long
computation chain -- its value will only become available during the build.}
of computing the dependency graph before the build starts.

To support dynamic dependencies, \Excel's calc engine~\cite{excel_recalc} is
significantly different from \Make. \Excel arranges the cells into a linear
sequence, called the \emph{calc chain}.  During the build, \Excel processes
cells in the calc-chain sequence, but if computing a cell \cmd{C} requires the
value of a cell \cmd{D} that has not yet been computed, \Excel \emph{aborts}
computation of \cmd{C}, moves \cmd{D} before \cmd{C} in the calc chain, and
resumes the build starting with \cmd{D}. When a build is complete, the resulting
calc chain respects all the dynamic dependencies of the spreadsheet. When an
input value (or formula) is changed, \Excel uses the final calc chain from the
\emph{previous} build as its starting point so that, in the common case where
changing an input value does not change dependencies, there are no aborts.
Notice that build always succeeds regardless of the initial calc chain (barring
truly circular dependencies); the calc chain is just an optimisation.
We refer to this algorithm as \emph{restarting}, and will discuss it
in more detail in~\S\ref{sec-implementation-excel}.

>>>>

Build systems automate the execution of repeatable tasks for individual users
and large organisations. In this section we explore the design space of build
systems, using four concrete examples: \Make~\cite{feldman1979make},
\Shake~\cite{mitchell2012shake}, \Bazel~\cite{bazel}, and
\Excel~\cite{advanced_excel}.
We have carefully chosen these four to illustrate the various axes on which
build systems differ; we discuss many other notable examples of build systems,
and their relationships, in~\S\ref{sec-related}.

@@

\noindent
There are two input cells \cmd{A1} and \cmd{A2}, and a single task that computes
the sum of their values, producing the result in cell \cmd{B1}. If either of
the inputs change, \Excel will recompute the result.

@@

\noindent
The formula in \cmd{B1} uses the \cmd{INDIRECT} function, which takes a string
and returns the value of the cell with that name.  The string evaluates to
\cmd{"A1"}, so \cmd{B1} evaluates to \cmd{10}. However, the dependencies of the
formula in \cmd{B1} are determined by the value of \cmd{C1}, so it is impossible
to compute the dependency graph before the build starts\footnote{In
this particular example one might say that the value of
\cmd{C1} is statically known, but imagine that it is the result of a long
computation chain -- its value will only become available during the build.}.

To support dynamic dependencies, \Excel's calc engine~\cite{excel_recalc} is
significantly different from \Make. \Excel arranges the cells into a linear
sequence, called the \emph{calc chain}.  During the build, \Excel processes
cells in the calc-chain sequence, but if computing a cell \cmd{C} requires the
value of a cell \cmd{D} that has not yet been computed, \Excel \emph{aborts}
computation of \cmd{C}, moves \cmd{D} before \cmd{C} in the calc chain, and
resumes the build starting with \cmd{D}. When a build is complete, the resulting
calc chain respects all the dynamic dependencies of the spreadsheet. When an
input value (or formula) is changed, \Excel uses the final calc chain from the
\emph{previous} build as its starting point so that, in the common case where
changing an input value does not change dependencies, there are no aborts.
Notice that build always succeeds regardless of the initial calc chain (barring
truly circular dependencies); the calc chain is just an optimisation.
We refer to this algorithm as \emph{restarting}, and discuss it
in more detail in~\S\ref{sec-implementation-excel}.
